%!TEX root = index.tex

\section{Flight Plan}

Every controlled aircraft has its flight plan (FP) that contains basic information about the flight. These information include aircraft identification, aircraft type, departure point, departure time, route of flight, destination, fuel on board and a few other. \cite[Chapter 5]{aim} The purpose of the flight plan is to give all relevant data about the aircraft to the air traffic controller in a standardized format so it can be used efficiently during the flight control. The flight plan is handed over from one control to another as the aircraft progresses from its departure airport to destination along the flight route.

The flight route is a sequence of fixes and routes the plane will fly through during its flight.

\subsection{Fix}
Fix is named point on Earth that aircraft use for their navigation. The position of fix can be determined by radio beacons or as GPS position for airplanes equipped with GPS receivers.

If aircraft has certain fix in its flight plan, it must fly over the position of the fix during the flight. This way the route of the aircraft is defined and automatic tools can be used to notify the pilot and air traffic controller if the plane deflects too far from the trajectory defined by its flight plan.

\subsection{Route}
Route is named sequence of fixes. There can be some additional restrictions added on the route, for example altitude at given fixes or type of aircraft that can use the route. The flight plan does not need to include the whole route, the plane can be ordered to follow only part of the route between two fixes.

\subsubsection{STAR}
Standard Terminal Arrival Route (STAR) is a special kind of route that is used to navigate airplanes approaching an airport. It links a significant fix with a point from witch the aircraft can safely land using instrument approach procedure. \cite{doc4444} This basically means the STAR leads the aircraft from its en-route flight to a touch-down on a runway.

\subsubsection{SID}
Standard Instrument Departure (SID) route works the other way around. It leads the aircraft from the runway to a specified significant fix from which the airplane can continue with its en-route flight.