%!TEX root = index.tex

\cleardoublepage 
\begin{thebibliography}{1}

\bibitem{growth}
ICAO.
\newblock {\em Strong Passenger Results and a Rebound for Freight Traffic in 2014.}
\newblock International Civil Aviation Organization, Montréal, 18 December 2014.
\newblock Available at: \url{http://www.icao.int/Newsroom/NewsDoc2014/COM.48.14.EN.pdf}

\bibitem{atm}
Eurocontrol.
\newblock {\em What is air traffic management} [online].
\newblock European Organisation for the Safety of Air Navigation [cited 2014-11-16].
\newblock Available at: \url{http://www.eurocontrol.int/articles/what-air-traffic-management}

\bibitem{nolan}
M.~S. Nolan.
\newblock {\em Fundamentals of air traffic control}.
\newblock Thomson--Brooks/Cole, Belmont, CA, 4th edition, 2004.

\bibitem{aim}
United States Department of Transportation, Federal Aviation Administration.
\newblock {\em Aeronautical Information Manual - Official Guide to Basic Flight Information and ATC Procedures}.
\newblock United States Department of Transportation, Federal Aviation Administration, 2014.
\newblock Available at: \url{https://www.faa.gov/air_traffic/publications/ATpubs/AIM/}.

\bibitem{doc4444}
ICAO.
\newblock {\em Doc 4444 – Air traffic management}.
\newblock International Civil Aviation Organization, Montréal, 15th edition, 2007.

\bibitem{doc8168}
ICAO.
\newblock {\em Doc 8168 – Aircraft Operations Volume I, Flight Procedures}.
\newblock International Civil Aviation Organization, Montréal, 5th edition, 2006.

\bibitem{annex11}
ICAO.
\newblock {\em Annex 11 to the Convention on International Civil Aviation: Air Traffic Services}.
\newblock International Civil Aviation Organization, 2001.

\bibitem{annex6}
ICAO.
\newblock {\em Annex 6 to the Convention on International Civil Aviation: Operation of Aircraft (Part I: International Commercial Air Transport — Aeroplanes)}.
\newblock International Civil Aviation Organization, 9th edition, 2010.

\bibitem{classes}
{\em Airspace class. in: Wikipedia: the free encyclopedia} [online].
\newblock San Francisco: Wikimedia foundation, 2001- [cited 2014-11-18].
\newblock Available at: \url{http://en.wikipedia.org/wiki/Airspace_class}

\bibitem{order7110}
United States Department of Transportation, Federal Aviation Administration.
\newblock {\em ORDER JO 7110.65U: Air Traffic Control}.
\newblock Available at: \url{http://www.faa.gov/air_traffic/publications/atpubs/ATC/index.htm}.

\bibitem{vor}
{\em VHF omnidirectional range. in: Wikipedia: the free encyclopedia} [online].
\newblock San Francisco: Wikimedia foundation, 2001- [cited 2014-11-26].
\newblock Available at: \url{http://en.wikipedia.org/wiki/VHF_omnidirectional_range}

\bibitem{agentfly-atm}
D. Šislák, P. Volf, D. Pavlíček and M. Pěchouček
\newblock {\em AGENTFLY: Multi-Agent Simulation of Air-Traffic Management}.
\newblock 2012.

\bibitem{agentfly-enroute}
D. Šišlák, P. Volf, M. Pěchouček, Ch. T. Cannon, D. N. Nguyen and W. C. Regli
\newblock {\em Multi-Agent Simulation of En-Route Human Air-Traffic Controller}.
\newblock 2012.

\bibitem{bada}
Eurocontrol.
\newblock {\em Base of Aircraft Data (BADA)} [online].
\newblock European Organisation for the Safety of Air Navigation [cited 2014-11-30].
\newblock Available at: \url{https://www.eurocontrol.int/services/bada}

\bibitem{bab}
J. Clausen.
\newblock {\em Branch and Bound Algorithms – Principles and Examples}.
\newblock 1999.
\newblock Available at: \url{http://www.imada.sdu.dk/Employees/jbj/heuristikker/TSPtext.pdf}

\bibitem{runway-throughput}
Tamas Kolos-Lakatos.
\newblock {\em The Influence of Runway Occupancy Time and Wake Vortex Separation Requirements on Runway Throughput}.
\newblock MIT International Center for Air Transportation (ICAT) Department of Aeronautics \& Astronautics Massachusetts Institute of Technology, 2013.

\bibitem{atlanta}
{\em Hartsfield–Jackson Atlanta International Airport. in: Wikipedia: the free encyclopedia} [online].
\newblock San Francisco: Wikimedia foundation, 2001- [cited 2014-12-03]. \\
\newblock Available~at: \url{http://en.wikipedia.org/wiki/Hartsfield\%E2\%80\%93Jackson_Atlanta_International_Airport}

\bibitem{atlanta-diagram}
United States Department of Transportation, Federal Aviation Administration.
\newblock {\em Airport Diagram – Hartsfield-Jackson Atlanta Intl. (ATL).}
\newblock 11 December 2014.
\newblock Available~at: \url{http://aeronav.faa.gov/d-tpp/1413/00026ad.pdf}

\bibitem{atlanta-tma}
United States Department of Transportation, Federal Aviation Administration.
\newblock {\em Document 2013-00287 – Amendment to Class B Airspace; Atlanta, GA.}
\newblock 9 January 2013.
\newblock Available~at: \url{https://www.federalregister.gov/articles/2013/01/09/2013-00287/amendment-to-class-b-airspace-atlanta-ga}

\bibitem{scheduling}
Susanne Albers. 
\newblock {\em Online scheduling.}
\newblock In Introduction to Scheduling, edited by Yves Robert and Frederic Vivien. 
\newblock Chapman and Hall/CRC Press, 57-84, 2009.
\newblock Available~at: \url{http://www14.in.tum.de/personen/albers/papers/sched_chapter.pdf}

\end{thebibliography}
\newpage